Manuscript_Github_Pchabaudi_pir

### 2021-03-15 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Reorder bits

---
title: "Manuscript_Github_Pchabaudi_pir_210105"
author: "Timothy Little"
date: "05/01/2021"
output: html_document
---

@@

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome.

@@

@@

## Figure 4 - Expression of major expressed cirs rMT/SBP side-by-side.

```{r pirs_side-by-side_figure4}

cir_plot <- twentyfour_ct_comb %>% 
  filter(Geneid %in% cir_housekeeping_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_info$subfam[match(Geneid, cir_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown'))) %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a'))

side_by_side_cir_plot <- function(time_vec = '14h',
                                  log_it = TRUE,
                                  limits = limits_cir,
                                  point_size = 1,
                                  text_size = 10,
                                  file_name_end = '',
                                  threshold_min = 100,
                                  threshold_max = Inf){

  cir_plot <- group_by(cir_plot, 
                       Geneid) %>% 
    filter(
      time == time_vec,
      !Geneid %in% cir_id | any(tpm >= threshold_min),
      !Geneid %in% cir_id | all(tpm <= threshold_max)
    )

  # cir_plot %>% group_by(Geneid) %>% summarise(n())
  if(log_it){
    cir_plot$tpm <- log2(1+cir_plot$tpm)
      limits = log2(1+limits)
      breaks <- round(seq.int(0, 
                              max(limits), 
                              length.out = round(max(limits)) + 1, 
                              digits = 0), 
                      digits = 0)
    }else{
      breaks <- seq.int(0, max(limits), length.out = 11)
    }

    cir_plot$subfam <- factor(
      cir_plot$subfam,
      levels = c(
        sort(unique(cir_plot$subfam)[!(unique(cir_plot$subfam) %in% 'ref')]),
        'ref'
      )
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      facet_grid(.~subfam, 
                 scales = "free_x",
                 space = "free_x") +
      theme_classic() +
      theme(
        axis.text.x = element_text(size = text_size,
                                   angle = 90)
      ) +
      ylab('log(TPM + 1)')+
      ggtitle(paste(unique(cir_plot$time)))

    g <- ggplot_gtable(ggplot_build(y))
    strip_t <- which(grepl('strip-t', g$layout$name))
    subfam_vec <- sort(unique(cir_plot$subfam))
    fills <- case_when(str_detect(subfam_vec, 
                                  pattern = 'a') ~ clade_col['ancestral'],
                       subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                       str_detect(subfam_vec, 
                                  pattern = 'ref') ~ 'white')
    k <- 1
    for (i in strip_t) {
      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
      k <- k+1
    }

    grid.newpage()
    grid.draw(g)

 lapply(
    sort(unique(cir_plot$time)),
    side_by_side_cir_plot,
    threshold_min = 100,
    threshold_max = Inf,
    point_size = 1,
    text_size = 10,
    file_name_end = '_Over100TPM'
  )

# Supp figure - Pir expression of AAPLs and ChAPLs.

```{r chrom_prop_bar_chart}
tpm_data_chrom <- tpm_data_twentyfour %>%
  mutate(stage = times_stages_table$time[match(stage,
                                               times_stages_table$r_stage_diff_count)]) %>%
  group_by(transmission,
           stage,
           chrom) %>%
  dplyr::summarise(total_tpm = sum(tpm))

y <- ggplot(data = tpm_data_chrom) +
  geom_bar_pattern(
    aes(
      x = stage,
      y = total_tpm, 
      fill = chrom,
      group = transmission,
      pattern = chrom
    ),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1,
    pattern_density = 0.05,
    pattern_spacing = 0.02,
    pattern_key_scale_factor = 0.5
  ) +
  scale_pattern_manual(
    values = rep(c('none', 
                   'circle'), 
                 each = length(c(cbPalette, 
                                 'red', 
                                 'white')), 
                 length.out = length(unique(tpm_data_chrom$chrom))),
    labels = unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_fill_manual(
    breaks = unique(tpm_data_chrom$chrom),
    values = rep_len(c(cbPalette, 
                       'red', 
                       'white'),
                     length.out = length(unique(tpm_data_chrom$chrom))),
    labels =  unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('proportion of '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~transmission, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_text(size = 10, 
                                 angle = 0),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=10, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

### 2D plot of PC1 v PC2 - Supp figure

@@

>>>>

---
title: "Manuscript_Github_Pchabaudi_pir"
author: "Timothy Little"
date: "12/03/2021"
output: html_document
---

@@

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.

@@

cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')
```

@@

### 2D plot of PC1 v PC2 - Supp figure 4A

@@

## Figure 4 and supplementary figure 7/8

```{r pirs_side-by-side_figure4_RMT-SBP-completelySeparate}

cir_plot <- twentyfour_ct_comb %>% 
  filter(Geneid %in% cir_housekeeping_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown'))) %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a')) %>% 
  mutate(
    subfam = interaction(
      transmission,
      forcats::fct_shift(
        relevel(
          as.factor(subfam),
          ref = 'ref')
      ),
      sep = '_',
      lex.order = TRUE
    )
  )
#Need to add the transmission name to the subfam so that they are ordered correctly.
# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.

transmissionSep_cir_plot <- function(time_vec = '14h',
                                     log_it = TRUE,
                                     limits = limits_cir,
                                     point_size = 1,
                                     text_size = 10,
                                     threshold_min = 100,
                                     threshold_max = Inf){

  cir_plot_filter <- group_by(cir_plot, 
                       Geneid) %>% 
    filter(
      time == time_vec,
      !Geneid %in% cir_id | any(tpm >= threshold_min),
      !Geneid %in% cir_id | all(tpm <= threshold_max)
    )
  #Filter for pirs expressed above or below the thresholds, have the OR so that the ref genes aren't removed.
  # e.g. Either 'Not a cir' OR 'Any TPM is above the minimum'.

  # cir_plot_filter %>% group_by(Geneid) %>% summarise(n())
  if(log_it){
    cir_plot_filter$tpm <- log2(1+cir_plot_filter$tpm)
    limits = log2(1+limits)
    breaks <- round(seq.int(0, 
                            max(limits), 
                            length.out = round(max(limits)) + 1, 
                            digits = 0), 
                    digits = 0)
  }else{
    breaks <- seq.int(0, max(limits), length.out = 11)
  }

  cir_plot_filter$Geneid <- factor(
    str_remove(cir_plot_filter$Geneid, 
               pattern = 'PCHAS_'),
    levels = str_remove(levels(cir_plot_filter$Geneid), 
                        pattern = 'PCHAS_')
  )

  lapply(unique(cir_plot_filter$transmission), 
         function(transmission_method){

           cir_plot_filter <- filter(cir_plot_filter, 
                              transmission %in% transmission_method)

           y <- ggplot(
             data = cir_plot_filter,
             mapping = aes(
               x = Geneid,
               y = tpm
             )
           ) +
             stat_summary(fun.y = median, 
                          fun.ymin = median, 
                          fun.ymax = median,
                          geom = "bar",
                          fill = 'lightblue',
                          width = 0.75, 
                          size = 0.25, 
                          position = 'dodge') +
             geom_point(stat = 'identity', 
                        position = position_dodge(width = 0.75),
                        size = point_size) +
             scale_y_continuous(breaks = breaks,
                                limits = limits) +
             facet_grid(.~subfam, 
                        scales = "free_x",
                        space = "free_x",
                        labeller = labeller(subfam = function(string) {
                          str_remove_all(string, 
                                         pattern = '(rMT|SBP)_')
                        })) +
             theme_classic() +
             ylab('log(TPM + 1)')+
             ggtitle(paste(unique(cir_plot_filter$time)))

           if(grepl(time_vec, pattern = '20|23')) {
             y <- y + theme(
               axis.text.x = element_text(size = text_size,
                                          angle = 90,
                                          vjust = 0.5)
             )
           } else {
             y <- y + theme(
               axis.text.x = element_blank(),
               axis.ticks.x = element_blank()
             ) 
           }

           g <- ggplot_gtable(ggplot_build(y))

           # g$widths[5] = 2*g$widths[5]

           strip_t <- which(grepl('strip-t', g$layout$name))
           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
           fills <- case_when(str_detect(subfam_vec, 
                                         pattern = 'a') ~ clade_col['ancestral'],
                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                              str_detect(subfam_vec, 
                                         pattern = 'ref') ~ 'white')
           k <- 1
           for (i in strip_t) {
             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
             k <- k+1
           }

           #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
           # manually make them a bit wider.

           short_panels <- which(str_detect(subfam_vec, pattern = 'a|S1'))

           for(i in short_panels){
             num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
             g$widths[num] = 1.75*g$widths[num]
             #1.75 times the width of the right panels.
             # print(num)
           }

           grid.newpage()
           grid.draw(g)
         })

lapply(
  sort(unique(cir_plot$time)),
  transmissionSep_cir_plot,
  threshold_min = 50,
  threshold_max = Inf,
  point_size = 1,
  text_size = 7.5
)
```

# Supp figure 9 - Pir expression of AAPLs and ChAPLs.
